---
title: 'Assignment 02 (Discussion): Determinant of a Matrix'
author: "Forhad Akbar"
date: "02/04/2020"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---

## **Load Packages**

```{r load-packages, warning = FALSE, message = FALSE}
library(knitr)
```

## **Problem Statement**  
The problem C23, selected page 353,  
Doing the computations by hand, find the determinant of the matrix below.

```{r matrix}
matr = matrix(c (1,3,2,4,1,3,1,0,1), ncol=3 , nrow=3, byrow=T)
 print(matr)
```
## **Solution**  
Determinant of the matrix above can be compute by hand as following

det (matr)  

= $\begin{vmatrix} 1 & 3 & 2 \\ 4 & 1 & 3 \\ 1 & 0 & 1 \end{vmatrix}$

= $1\begin{vmatrix} 1 & 3 \\ 0 & 1 \end{vmatrix}-3\begin{vmatrix} 4 & 3 \\ 1 & 1 \end{vmatrix}+2\begin{vmatrix} 4 & 1 \\ 1 & 0 \end{vmatrix}$

= 1(1X1 - 3X0) - 3(4X1 - 3X1) + 2(4X0 -1X1)  
= 1(1) - 3(1) + 2(-1)  
= 1 - 3 - 2  
= -4  

Above we used  
$Determinant\quad of\quad Matrices\quad of\quad Size\quad Two\\ suppose\quad that\quad A\quad =\quad \begin{vmatrix} a & b \\ c & d \end{vmatrix}\quad then\quad det(A)\quad =\quad ad\quad -\quad bc$

We can verify the result using R

```{r verify-result}
det(matr)
```
