---
title: 'Assignment 12 (Discussion):'
author: "Forhad Akbar"
date: "04/15/2020"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---

## **Load Packages**

```{r load-packages, warning = FALSE, message = FALSE}
library(corrplot)
library(dplyr)
library(knitr)
```

## **Problem Statement**  

Using R, build a regression model for data that interests you. Conduct residual analysis. Was the linear model appropriate? Why or why not?

### **Solution**

The data i choose is the heart data from UCI machine learning reprository
https://archive.ics.uci.edu/ml/index.php

Read Data

```{r}
heart <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/processed.cleveland.data",header=FALSE,sep=",",na.strings = '?')
colnames(heart) <- c( "age", "sex", "cp", "trestbps", "chol","fbs", "restecg","thalach","exang", "oldpeak","slope", "ca", "thal", "outcome")
```


```{r}
heart<- heart[complete.cases(heart),] #Store only Data without missing value
```

correlation plot

```{r}
data_frame <- heart %>% select(age,trestbps,chol,thalach,oldpeak)
M <- cor(data_frame)

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
          insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
)
```

### *Multiple Linear Regression:*

```{r}

#lm(formula = Y(dependent) ~ X1(independent) + X2 + X3 + ..., data=table_name_optional)
  
attach(heart)

MLR <- lm(formula = age ~ trestbps + chol + thalach + oldpeak, data=heart)

```

```{r}
summary(MLR) #compelete summary of the regression analysis
```

```{r}
confint(MLR)
```

## *Residual analysis*

```{r}
plot(fitted(MLR), resid(MLR))
```

```{r}
qqnorm((resid(MLR)))
qqline((resid(MLR)))
```

Based on the residual analysis, I would conclude that the linear model is appropriate. From the plots you can see that the points follow a straight line. This tells us that the residuals are normally distributed.