---
title: 'DATA 605: Homework 04'
author: "Forhad Akbar"
date: "02/20/2020"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---
## **Load Packages**

```{r load-packages, warning = FALSE, message = FALSE}
library(knitr)
library(pracma)
```

## **1. Problem set 1**  

In this problem, we'll verify using R that SVD and Eigenvalues are related as worked out in the weekly module. Given a 3 x 2 matrix A  
<center>
$A\quad =\quad \begin{bmatrix} 1 & 2 & 3 \\ -1 & 0 & 4 \end{bmatrix}$
</center>
write code in R to compute $X=A{ A }^{ T }$ and $Y={ A }^{ T }A$. Then, compute the eigenvalues and eigenvectors of X and Y using the built-in commans in R. Then, compute the left-singular, singular values, and right-singular vectors of A using the svd command. Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y. In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A.   
Your code should compute all these vectors and scalars and store them in variables. Please add enough comments in your code to show me how to interpret your steps.

### Write matrix A
```{r}
A <- matrix(c(1,2,3,-1,0,4), 2, byrow=T)
# Display matrix A
A
```


### Compute X and Y

```{r}
X <- A %*% t(A)
# Display X
X
```

```{r}
Y <- t(A) %*% A
# Display Y
Y
```

### Compute Eigenvectors & Eigenvalues of X and Y

```{r}
eX <- eigen(X)
# Display eX
eX
```

```{r}
eY <- eigen(Y)
# Display eY
eY
```

### Compute Singular Values 

```{r}
svdA <- svd(A)
# Display svdA
svdA
```

### Compare values

```{r}
comp1 <- cbind(svdA$u, eX$vectors)
colnames(comp1) <- c('SVDu1', 'SVDu2', 'EVX1=u1', 'EVX2=u2')
knitr::kable(comp1)
comp2 <- cbind(svdA$v, eY$vectors)
comp2 <- comp2[,1:4]
colnames(comp2) <- c('SVDv1', 'SVDv2', 'EVY1=v1', 'EVY2=v2')
knitr::kable(comp2)
```


We can see from the table that the singular vectors obtained using the SVD command are the same as the eigenvectors of X and Y. The only difference is whether they are positive or negative, which is due to them being declared in unit vector form.   The ratios are otherwise consistent. 

# Problem 2

Using the procedure outlined in section 1 of the weekly handout, write a function to compute the inverse of a well-conditioned full-rank square matrix using co-factors. In order to compute the co-factors, you may use built-in commands to compute the determinant.

Your function should have the following signature:  
$B = myinverse(A)$  
where A is a matrix and B is its inverse and $A \times B = I$. The off-diagonal elements of I should be close to zero, if not zero. Likewise, the diagonal elements should be close to 1, if not 1. Small numerical precision errors are acceptable but the function myinverse should be correct and must use co-factors and determinant of A to compute the inverse.

### Setup MyInverse Function

```{r}
myinverse <- function(A, debug=FALSE){
    CoF <- diag(ncol<-nrow(A))
    for (i in 1:nrow(A)) { 
        for (j in 1:ncol(A)){ 
            CoF[i,j]=((-1)^(i+j))*det(A[-i,-j])
        }
        if (debug==TRUE){
            print(sprintf('CoF_(%s,%s):', i, j))
            print(CoF)            
            }
    }
    return(t(CoF)/det(A))
}
```

### Testing

```{r}
# T1
A <- matrix(c(1,2,3,2,1,3,1,1,3),nrow=3)
Ainv <- myinverse(A, debug=TRUE)
comp3 <- cbind(Ainv, solve(A))
colnames(comp3) <- c('myI1', 'myI2', 'myI3', 
                     'solve1', 'solve2', 'solve3')
knitr::kable(comp3)
# T2
B <- matrix(c(18, 8, 6, 8, 6, 2, 6, 2, 4), 3, byrow=TRUE)
Binv <- myinverse(B, debug=FALSE)
comp4 <- cbind(Binv, solve(B))
colnames(comp4) <- c('myI1', 'myI2', 'myI3', 
                     'solve1', 'solve2', 'solve3')
knitr::kable(comp4)
```
