---
title: 'Assignment 03 (Discussion): Chapter E'
author: "Forhad Akbar"
date: "02/11/2020"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---

## **Load Packages**

```{r load-packages, warning = FALSE, message = FALSE}
library(knitr)
library(pracma)
```

## **Problem Statement**  
The problem C11, selected page 388,  
Find the characteristic polynomial of the matrix.

```{r matrix}
A = matrix(c (3,2,1,0,1,1,1,2,0), ncol=3 , nrow=3, byrow=T)
print(A)
```

## **Solution using R** 

```{r}
# Calculate characteristic polynomial using pracma library
charpoly(A)
```

Which gives us the characteristic polynomial as

${ PA(x)\quad =\quad λ }^{ 3 }-4{ λ }^{ 2 }+5$

## **Solution by hand**  

Let’s find the characteristic polynomial of a 3x3 matrix. Let

$A=\begin{vmatrix} 3 & 2 & 1 \\ 0 & 1 & 1 \\ 1 & 2 & 0 \end{vmatrix}$

To get the eigenvalues, we need to solve

$det(\begin{vmatrix} λ & 0 & 0 \\ 0 & λ & 0 \\ 0 & 0 & λ \end{vmatrix}-\begin{vmatrix} 3 & 2 & 1 \\ 0 & 1 & 1 \\ 1 & 2 & 0 \end{vmatrix})\quad =\quad 0$

That is

$det(\begin{vmatrix} λ-3 & -2 & -1 \\ 0 & λ-1 & -1 \\ -1 & -2 & λ \end{vmatrix})\quad =\quad 0$

Solving determinant  
$(λ-3)\begin{vmatrix} λ-1 & -1 \\ -2 & λ \end{vmatrix}-(-2)\begin{vmatrix} 0 & -1 \\ -1 & λ \end{vmatrix}+(-1)\begin{vmatrix} 0 & λ-1 \\ -1 & -2 \end{vmatrix}\quad =\quad 0$

$(λ-3)\left[ ({ λ }^{ 2 }-λ)-2 \right] +2(λ-1)+2+(-λ+1)\quad =\quad 0\\ { λ }^{ 3 }-{ λ }^{ 2 }-2λ-3{ λ }^{ 2 }+3{ λ }+6+2{ λ }-2-{ λ }+1\quad =\quad 0\\ { λ }^{ 3 }-4{ λ }^{ 2 }+5\quad =\quad 0$

Which gives us the characteristic polynomial as

${ PA(x)\quad =\quad λ }^{ 3 }-4{ λ }^{ 2 }+5$
