---
title: 'DATA 605: Homework 03'
author: "Forhad Akbar"
date: "02/12/2020"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---
## **Load Packages**

```{r load-packages, warning = FALSE, message = FALSE}
library(knitr)
library(pracma)
```

## **1. Problem set 1**  

1. What is the rank of the matrix A?

$A\quad =\quad \begin{vmatrix} 1 & 2 & 3 & 4 \\ -1 & 0 & 1 & 3 \\ 0 & 1 & -2 & 1 \\ 5 & 4 & -2 & -3 \end{vmatrix}$

The set of pivot columns of any reduced row echelon form matrix is known as Rank.

Transform matrix A to reduced echelon form, which is:

$A\quad =\quad \begin{vmatrix} 1 & 0 & 0 & 0 \\ 0 & 1 & 0 & 0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 1 \end{vmatrix}$

Looking at the resultant matrix, the rank is 4.

Confirming rank using R

```{r}
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow=4, byrow=TRUE)

# matrix A
A
```

```{r}
# RREF of A
library(pracma)
rref(A)
```

There are 4 pivot rows, so the rank of A is 4. Double-check with R.
```{r}
# Rank of A
Rank(A)
```

2. Given an mxn matrix where m > n, what can be the maximum rank? The mini-mum rank, assuming that the matrix is non-zero?  

Since rank of a matrix is defined as the number of linearly independent column vectors and is equal to the number of linearly independent row vectors, given an m×n matrix with m>n, the maximum rank is the lower value ${ rank }_{ max }\quad =\quad n$. Assuming a non-zero matrix, the rank should be at least 1, so ${ rank }_{ min }\quad =\quad 1$.

3. What is the rank of matrix B?

$B\quad =\quad \begin{vmatrix} 1 & 2 & 1 \\ 3 & 6 & 3 \\ 2 & 4 & 2 \end{vmatrix}$

We can see that row 2 and row 3 are multiples of row 1, so rows 1 and 2 and 1 and 3 are linearly dependent. There is only 1 linearly independent row, so $rank(B)\quad =\quad 1$

Confirming rank using R
```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow=3, byrow=TRUE)

# matrix B
B
```

```{r}
# Rank of B
Rank(B)
```

## **2. Problem set 2**
Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

$A\quad =\quad \begin{vmatrix} 1 & 2 & 3 \\ 0 & 4 & 5 \\ 0 & 0 & 6 \end{vmatrix}$

Since, A is a triangular matrix, its eigenvalues are values on the diagonal, so λ1=1, λ2=4 and λ3=6. But i am going to show the work as following

Step 1 - Calculate the eigenvalues of A

$det(A-λI)\quad =\quad 0$  

$det(\begin{vmatrix} 1 & 2 & 3 \\ 0 & 4 & 5 \\ 0 & 0 & 6 \end{vmatrix}-\begin{vmatrix} λ & 0 & 0 \\ 0 & λ & 0 \\ 0 & 0 & 0 \end{vmatrix})\quad =\quad 0$  

$det(\begin{vmatrix} 1-λ & 2 & 3 \\ 0 & 4-λ & 5 \\ 0 & 0 & 6-λ \end{vmatrix})\quad =\quad 0$  

$(1-λ)\quad \begin{vmatrix} 4-λ & 5 \\ 0 & 6-λ \end{vmatrix}\quad =\quad 0$  

$(1−λ)((4−λ)(6−λ)\quad =\quad 0$  

So,The characteristic polynomial is  

$P(λ)=(1−λ)(4−λ)(6−λ)\quad or\quad P(λ)={ λ }^{ 3 }-11{ λ }^{ 2 }+34λ-24$  

We determine that the eigenvalues of A, given by λ, are 1, 4, 6.  

We can use R to confirm the result
```{r}
#define matrix
A = matrix(c(1,0,0, 2,4,0, 3,5,6),ncol = 3)

cp = charpoly(A)

#Characteristic polynomial
cp
```
```{r}
ev = eigen(A)
ev$values
```

Step 2 - Calculate the eigenvectors of A using λ  

Using λ1 = 1:  

$A−λI\quad =\quad \begin{vmatrix} 1−1 & 2 & 3 \\ 0 & 4-1 & 5 \\ 0 & 0 & 6-1 \end{vmatrix}\quad =\quad \begin{vmatrix} 0 & 2 & 3 \\ 0 & 3 & 5 \\ 0 & 0 & 5 \end{vmatrix}$  

Using Gaussian Elimination on the resulting augmented matrix:  
$\begin{vmatrix} 0 & 2 & 3 & 0 \\ 0 & 3 & 5 & 0 \\ 0 & 0 & 5 & 0 \end{vmatrix}$  
$R1=1/2(R1)\quad =\quad \begin{vmatrix} 0 & 1 & 3/2 & 0 \\ 0 & 3 & 5 & 0 \\ 0 & 0 & 5 & 0 \end{vmatrix}$  
$R2\quad =\quad −3(R1)\quad =\quad \begin{vmatrix} 0 & 1 & 3/2 & 0 \\ 0 & 0 & 1/2 & 0 \\ 0 & 0 & 5 & 0 \end{vmatrix}$  
$R2\quad =\quad 2(R2)\quad =\quad \begin{vmatrix} 0 & 1 & 3/2 & 0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 5 & 0 \end{vmatrix}$  
$R3\quad =\quad −5(R2)=\quad \begin{vmatrix} 0 & 1 & 3/2 & 0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  
$R1\quad =\quad R1−3/2(R2)\quad =\quad \begin{vmatrix} 0 & 1 & 0 & 0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  

For the eigenvalue λ1 = 1, the eigenvector is:  

${ v }_{ 1 }=\quad \begin{pmatrix} { x }_{ 1 } \\ 0 \\ 0 \end{pmatrix}\quad =\quad \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix}$  
----------------------------------------------------
Using λ2 = 4:

$A−λI\quad =\quad \begin{vmatrix} 1−4 & 2 & 3 \\ 0 & 4-4 & 5 \\ 0 & 0 & 6-4 \end{vmatrix}\quad =\quad \begin{vmatrix} -3 & 2 & 3 \\ 0 & 0 & 5 \\ 0 & 0 & 2 \end{vmatrix}$  

Using Gaussian Elimination on the resulting augmented matrix:

$\begin{vmatrix} -3 & 2 & 3 & 0 \\ 0 & 0 & 5 & 0 \\ 0 & 0 & 2 & 0 \end{vmatrix}$  
$R1\quad =\quad −1/3(R1)\quad =\quad \begin{vmatrix} 1 & -2/3 & -1 & 0 \\ 0 & 0 & 5 & 0 \\ 0 & 0 & 2 & 0 \end{vmatrix}$  
$R2\quad =\quad (R1)/5)\quad =\quad \begin{vmatrix} 1 & -2/3 & -1 & 0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 2 & 0 \end{vmatrix}$  
$R3\quad =\quad R3−3(R2)\quad =\quad \begin{vmatrix} 1 & -2/3 & -1 & 0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  
$R1\quad =\quad R1+R2\quad =\quad \begin{vmatrix} 1 & -2/3 & 0 & 0 \\ 0 & 0 & 1 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  
Solving the system:
${ x }_{ 1 }−2/3{ x }_{ 2 }\quad =\quad 0;\quad { x }_{ 1 }\quad =\quad 2/3{ x }_{ 2 }\\ { x }_{ 3 }\quad =\quad 0$  

For the eigenvalue λ2 = 4, the eigenvector is:  

${ v }_{ 2 }\quad =\quad \begin{pmatrix} 2/3{ x }_{ 2 } \\ { x }_{ 2 } \\ 0 \end{pmatrix}\quad =\quad \begin{pmatrix} 2 \\ 3 \\ 0 \end{pmatrix}$  
----------------------------------------------------
Using λ3 = 6:

$A−λI\quad =\quad \begin{vmatrix} 1−6 & 2 & 3 \\ 0 & 4-6 & 5 \\ 0 & 0 & 6-6 \end{vmatrix}\quad =\quad \begin{vmatrix} -5 & 2 & 3 \\ 0 & -2 & 5 \\ 0 & 0 & 0 \end{vmatrix}$  
Using Gaussian Elimination on the resulting augmented matrix:  
$\begin{vmatrix} -5 & 2 & 3 & 0 \\ 0 & -2 & 5 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  
$R1\quad =\quad −1/5(R1)\quad =\quad \begin{vmatrix} 1 & -2/5 & -3/5 & 0 \\ 0 & -2 & 5 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  
$R2\quad =\quad −1/2(R2)\quad =\quad \begin{vmatrix} 1 & -2/5 & -3/5 & 0 \\ 0 & 1 & -5/2 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  
$R1=R1−2/5(R2)\quad =\quad \begin{vmatrix} 1 & 0 & -8/5 & 0 \\ 0 & 1 & -5/2 & 0 \\ 0 & 0 & 0 & 0 \end{vmatrix}$  
Solving the system:

${ x }_{ 1 }+0{ x }_{ 2 }-8/5{ x }_{ 3 }\quad =\quad 0\\ { 0x }_{ 1 }+{ x }_{ 2 }-5/2{ x }_{ 3 }\quad =\quad 0$  
${ x }_{ 1 }=8/5{ x }_{ 3 }\\ { x }_{ 2 }=5/2{ x }_{ 3 }$  

For the eigenvalue λ3 = 6, the eigenvector is:

${ v }_{ 3 }\quad =\quad \begin{pmatrix} 8/5{ x }_{ 3 } \\ 5/2{ x }_{ 3 } \\ 10 \end{pmatrix}\quad =\quad \begin{pmatrix} 16 \\ 25 \\ 10 \end{pmatrix}$  
----------------------------------------------------