---
title: 'Assignment 05 (Discussion): Chapters 2'
author: "Forhad Akbar"
date: "02/26/2020"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---

## **Load Packages**

```{r load-packages, warning = FALSE, message = FALSE}
library(knitr)
```

## **Problem Statement**  
The problem 6, selected page 72, from "Introduction to Probability" book

Assume that a new light bulb will burn out after t hours, where t is chosen from [0,∞) with an exponential density
$$
f(t)=\lambda { e }^{ -\lambda t }
$$

In this context, λ is often called the failure rate of the bulb.

(a) Assume that λ = 0.01, and find the probability that the bulb will not burn out before T hours. This probability is often called the reliability of the bulb.  

(b) For what T is the reliability of the bulb = 1/2?

## **Solution (a)**  

If X is a random variable, then we essentially want to find the probability of X less than or equal to T, where T is the time of failure. Hence we have the following:

The probability that the bulb will burn our before T hours is 
$$
P(x\le T)
$$

The probability that the bulb will not burn out after T hours is simply the compliment 
$$
1-P(x\le T)
$$

Since we are dealing with a continous exponential function of time, we should consider using an integral with integration limits 0 to T and integrand f(t)

$$
1-\int _{ 0 }^{ T }{ \lambda { e }^{ -\lambda t } } dt
$$
We are given a value of lambda= 0.01
$$
1-\int _{ 0 }^{ T }{ 0.01 { e }^{ -0.01 t } } dt\\
1-0.01\int _{ 0 }^{ T }{  { e }^{ -0.01 t } } dt
$$
The technique at hand is u-substitution (reverse chain rule)
https://www.khanacademy.org/math/ap-calculus-ab/ab-integration-new/ab-6-9/a/review-applying-u-substitution
$$
u=-0.01t\\
du=-0.01dt\\
 \frac { -du }{0.01  } =dt
$$
$$
1-0.01\int _{ 0 }^{ T }{  { e }^{ u } } \frac { -du }{0.01  }\\
1+\int _{ 0 }^{ T }{  { e }^{ u } } du\\
$$
Transform upper and lower limits using our u-sub 
$$
Upper: u=-0.01t=-0.01T\\
lower: u=-0.01t=-0.01(0)=0
$$
Replace newlimits and evaluate in terms of u 
$$
1+\int _{ 0 }^{-0.01 T }{  { e }^{ u } } du\\
$$
Apply fundemental theorem of calculus for definite integrals (part 2 of theorem)
http://mathworld.wolfram.com/FundamentalTheoremsofCalculus.html
$$
1+(e^{-0.01T}-e^{0})=\\
1+e^{-0.01T}-1=\\
=e^{-0.01T}
$$

## **Solution (b)**  

We need to perform some algebra and solve for T using our derived probability. We need to solve for T  
We need to solve exponential equations 
http://www.sosmath.com/algebra/logs/log4/log46/log46.html
$$
\frac { 1 }{2 }=e^{-0.01T}
$$
$$
ln(1/2)=ln(e^{-0.01T})\\
ln(1/2)=-0.01T\\
\frac { ln(1/2) }{-0.01 }=T
$$
Using quotient to subtraction property of ln
$$
\frac { ln(1)-ln(2) }{-0.01 }=T\\
\frac { 0-ln(2) }{-0.01 }=T\\
\frac { -ln(2) }{-0.01 }=T\\
\frac { ln(2) }{0.01 }=T
$$

```{r}
log(2)/.01
```
If you calculate these values in a calculator or in R, then T is roughly 70, hence 70 hours. 
