---
title: 'DATA 605: Homework 05'
author: "Forhad Akbar"
date: "02/28/2020"
output:
  prettydoc::html_pretty:
    pdf_document: default
    theme: cayman
  html_document:
    df_print: paged
  pdf_document:
    number_sections: no
---
## **Load Packages**

```{r load-packages, warning = FALSE, message = FALSE}
library(knitr)
```

## **Problem set**  

Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square.

## **Proof:**

```{r}
# generate random numbers from a uniform distribution
set.seed(NULL)
B <- runif(10000, min = 0, max = 1)  

set.seed(NULL)
C <- runif(10000, min = 0, max = 1)   

df_b<- as.data.frame(B)

# check whether the probability of B is between 0 and 1

# firsr create a series with all the values as False
sum_p <- rep(F,times=length(B))

# then calculate the pribability of B
for (i in 1:length(B)){
  x=1
  for (j in 1:length(B)){
    if (j != i && B[i] == B[j]){
      x <- x + 1
    }
  }
  p <- x/length(B)
  # if the probability is between 0 and 1 inclusive, the value in sum_p will change to True correspondingly 
  sum_p[i] <- (0 <= p && p<= 1) 
}

condition_1 <- !any(sum_p == F)

print(paste("Is all the probabilities of B is between 0 and 1 inclusive? ",condition_1 ))
```

```{r}
print(paste("Similarly, all the probabilities of B is between 0 and 1 inclusive "))
```

```{r}
# sum the probability of B and denote as sigma
sigma_b <- 0
for (i in 1:length(B)){
  x=1
  for (j in 1:length(B)){
    if (j != i && B[i] == B[j]){
      x <- x + 1
    }
  }
  p <- x/length(B)
  sigma_b <- sigma_b + p 
}

sigma_b
```

```{r}
sigma_b <- round(sigma_b)

print(paste("The sum of the probabilities of the outcomes of B is",sigma_b))
```

```{r}
print(paste("Similarly,The sum of the probabilities of the outcomes of C is",sigma_b))
```

For both B or C, all the probabilities is between 0 and 1 inclusive.  

For both B or C,the sum of the probabilities of the outcomes is 1.  

This proof that B and C are proper probability distributions.


Find the probability that  
(a) $B+C<1/2$
```{r}
j = 0
for(i in 1:length(B)){
  if(B[i]+C[i] < 0.5){
    j = j+1
  }
}
print(paste("The Probabilty B+C < 1/2 =", j/length(B)))
```

The probabilty is 12.5%. For this to be true B and C must be <12, which is 0.5∗0.5=0.25 and C or B must be <12−[Compliment] the range where is can be true is [0,0.5]. The overall probabilty is 0.25∗0.5=0.125 which is what we see in the simulation.

(b) $BC<1/2$
```{r}
j = 0
for(i in 1:length(B)){
  if(B[i]*C[i] < 0.5){
    j = j+1
  }
}
print(paste("The Probabilty BC < 1/2 =", j/length(B)))
```

(c) $|B−C|<1/2$

```{r}
j = 0
for(i in 1:length(B)){
  if(abs(B[i]-C[i]) < 0.5){
    j = j+1
  }
}
print(paste("The Probabilty |B-C| < 1/2 =", j/length(B)))
```

(d) $max{B,C}<1/2$

```{r}
j = 0
for(i in 1:length(B)){
  if(max(c(B[i],C[i])) < 0.5){
    j = j+1
  }
}
print(paste("The Probabilty max(B,C) < 1/2 =", j/length(B)))
```

(e) $min{B,C}<1/2$

```{r}
j = 0
for(i in 1:length(B)){
  if(min(c(B[i],C[i])) < 0.5){
    j = j+1
  }
}
print(paste("The Probabilty min(B,C) < 1/2 =", j/length(B)))
```